### 2019-12-03 rmds/linear_regression.Rmd
add rmds folder

---
title: "Linear Regression (TMAs)"
author: "Raven McKnight"
date: "12/3/2019"
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: paper
    code_folding: hide
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tigris)
library(ggplot2)
library(leaps)

options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

counties <- c("Anoka", "Carver", "Dakota", "Hennepin", "Ramsey", "Scott", "Washington")
bgs <- block_groups("MN", counties, 2016)

# should find better way to read data from rmd folder
lm_dat <- readRDS('/Users/mcknigri/Documents/honors/honors-project/data/modeling-dat/lm_dat_logged.RDS')
```

# Introduction

The simplest model we can fit is a linear regression. This model was inspired by Metro Transit's Transit Market Areas. Here, we'll let $Y_i$ equal the log of average number of weekday boardings per stop in census block group $i$ in 2017. We'll use four basic predictors, inspired by the original Transit Market Areas: population density, employment density, percent of households with no vehicle, and a walkability metric derived from walk isochrones. We also include number of stops made in block group $i$ as a control for level of service.

# The Model

```{r}
lm1 <- lm(log_daily_boards ~ pop_density + emp_density + walkability + perc_no_veh + log_daily_stops, lm_dat, na.action = na.exclude)

summary(lm1)
```

I'm guessing population density is insignificant because it's so correlated with employment.

# Results
## Transit Market Areas

```{r}
lm_dat[, index := -3.79 + 0.06*pop_density + 0.22*emp_density + 0.21*walkability + 0.19*perc_no_veh + 1.18*log_daily_stops]

lm_dat[, predicted_boardings := exp(index)]

lm_dat_sf <- dplyr::left_join(bgs, lm_dat)
```

We can use the model above to define updated Transit Market Areas based on the number of boardings a block group can support, rather than an arbitary index value. The distribution of predicted boardings per stop is below:

```{r}
ggplot(lm_dat, aes(x=predicted_boardings)) + 
  geom_histogram() +
  theme_minimal() +
  labs(title = "Predicted Boardings")
```

```{r}
ggplot(lm_dat, aes(x=daily_boards, y = predicted_boardings)) +
  geom_point() +
  geom_smooth()

ggplot(lm_dat[daily_boards < 200], aes(x=daily_boards, y = predicted_boardings)) +
  geom_point() +
  geom_smooth()
```

There are several ways we can split this distribution to achieve 5 distinct market areas. However, a map of the continuous estimate is also informative:

```{r}
ggplot(lm_dat_sf %>% dplyr::filter(!is.na(predicted_boardings))) +
  geom_sf(aes(fill = predicted_boardings), lwd = 0) +
  scale_fill_viridis_c() +
  theme_minimal()
```

Here we can see the "bullseye" pattern we might expect

## Residuals