### 2019-12-03 rmds/spatial_mod.Rmd

---
title: "Spatial Model"
author: "Raven McKnight"
date: "12/3/2019"
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: paper
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction