### 2019-12-13 rmds/spatial_mod.Rmd

---
title: "Spatial Model"
author: "Raven McKnight"
date: "12/3/2019"
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: paper
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(INLA)
library(rstan)
library(shinystan)
library(data.table)
library(ggplot2)
library(rgdal)
library(sf)
library(tigris)
library(spdep)
library(bayesplot)

options(mc.cores = parallel::detectCores())
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

mod_dat <- readRDS('/Users/raven/Documents/honors/honors-project/data/modeling-dat/p_dat_scaled.RDS')
setDT(mod_dat)

counties <- c("Anoka", "Carver", "Dakota", "Hennepin", "Ramsey", "Scott", "Washington")
bgs <- block_groups("MN", counties)

spat_dat <- dplyr::left_join(bgs, mod_dat)
spat_dat <- spat_dat %>%
  dplyr::filter(!is.na(daily_boards) & !is.na(walkability) & !is.na(perc_no_veh) & !is.na(pop_density) & !is.na(emp_density))
```

# Introduction

We can improve the Poisson regression by modeling the residuals. As we saw, there is some spatial autocorrelation in the residuals. Additionally, the distribution of $Y_i$ is overdispersed, meaning the variance is greater than we would expect from a Poisson distribution. Therefore, we'll add *two* parameters to account for both pure, uncorrelated overdispersion ($\theta$) and spatially autocorrelated variance ($\phi$).

## The Model
For areal unit $i$, the spatial model can be defined

Let $Y_i$ represent the average number of daily boardings in census block group $i$ (rounded to an integer). Then the spatial model can be defined

\begin{align}
Y_i|\theta_i & \sim \text{Poisson}(E_i\lambda_i) \\
\eta_i = \text{log}(\lambda_i) &= \mu + x\beta + ((\sqrt{\rho / s})\phi + (\sqrt{1-\rho})\theta)\sigma  \\
\phi &\sim N(0, Q^{-1}) \\
\theta &\sim N(0, n) \\
\rho & \sim \text{beta}(1/2, 1/2) \\
\sigma &\sim \text{exponential }

\end{align}

where

* E is an offset term, sometimes also called the baseline or "expected value" of Y (same as Poisson regression)
* $\lambda_i$ is the relative "risk", and $\eta_i$ is the log relative risk
* $\mu$ is the overall "risk" level, an intercept term
* $x\beta$ are covariates which vary spatially
* $\phi$ is the intrinsic conditional autoregressive (ICAR) model which accounts for spatially structured errors. In this case, we are using the BYM2 model
* $\theta$ accounts for pure overdispersion. This parameter is the complement to $\phi$ and accounts for non-spatial errors. 
* $\rho$ is a mixing parameter. It determines how much variance comes from spatially correlated versus independent error terms. The introduction of $\rho$ is a key difference between the BYM and BYM2 models. 
* $s$ is a scaling factor such that $Var(\phi_i) \approx 1$. We set this variance so that $\sigma$ is the actual standard deviation of the errors.
* $n$ is the number of fully connected subgraphs in the neighborhood structure. $n=1$ for study areas with no regions with 0 neighbors. 
* $\sigma \geq 0$ is the overall standard deviation for combined error terms

# Results

## One covariate, Ramsey County
```{r}
spat_dat <- spat_dat %>%
  dplyr::filter(COUNTYFP == 123)

y = spat_dat$daily_boards
E = spat_dat$daily_stops

K = 1
x = matrix(ncol = 1, c(spat_dat$emp_density))

nb <- poly2nb(spat_dat)

source("nb_helper_funs.R")
nbs <- nb2graph(nb)
N <- nbs$N
node1 <- nbs$node1
node2 <- nbs$node2
N_edges <- nbs$N_edges
scaling_factor <- scale_nb_components(nb)[1]

standat1 <- list(N = N, N_edges = N_edges, node1 = node1, node2 = node2, y = y, E = E, scaling_factor = scaling_factor, K = K, x = x)

# bym2 <- "/Users/raven/Documents/honors/honors-project/stan/bym2.stan"
# 
# #bym2_fit1 <- stan(bym2, data = standat1, warmup = 5000, iter = 6000, verbose = T)
# #saveRDS(bym2_fit1, "/Users/raven/Documents/honors/honors-project/fits/bym2_fit1.RDS")
# 
# bym2_fit1 <- readRDS("/Users/raven/Documents/honors/honors-project/fits/bym2_fit1.RDS")
```

<!-- ```{r} -->
<!-- print(bym2_fit1, pars=c("beta0", "betas[1]", "rho", "sigma", "logit_rho", "mu[5]", "phi[5]", "theta[5]"), probs=c(0.025, 0.5, 0.975)) -->
<!-- #launch_shinystan(bym2_fit1) -->
<!-- ``` -->

<!-- I'm not sure if the problem with this is that I didn't run it enough, it's parameterized poorly, etc. Actually only took ~10 minutes to run.  -->

# All four covariates, Ramsey County

```{r}
x2 <- matrix(ncol = 4, c(spat_dat$pop_density, spat_dat$emp_density, spat_dat$perc_no_veh, spat_dat$walkability))
K2 <- 4

standat2 <- list(N = N, N_edges = N_edges, node1 = node1, node2 = node2, y = y, E = E, scaling_factor = scaling_factor, K = K2, x = x2)

#bym2_fit2 <- stan(bym2, data = standat2, warmup = 2000, iter = 3000, thin = 1, verbose = T)

#saveRDS(bym2_fit2, "/Users/raven/Documents/honors/honors-project/fits/bym2_fit2.RDS")

bym2_fit2 <- readRDS("/Users/raven/Documents/honors/honors-project/fits/bym2_fit2.RDS")
```

```{r}
launch_shinystan(bym2_fit2)
print(bym2_fit2, pars=c("beta0", "betas[1]", "betas[2]", "betas[3]", "betas[4]", "rho", "sigma", "logit_rho", "mu[5]", "phi[5]", "theta[5]"), probs=c(0.025, 0.5, 0.975))
```

Some of these terms haven't fully converged but I've only run the chains for 2000 iterations. Using `shinystan`, I can see that there's a lot of autocorrelation in the chains. I think if there are no other red flags I can run this for many more iterations for a good result.

```{r}
posteriorbym2 <- rstan::extract(bym2_fit2)
yrepbym2 <- matrix(ncol = 357, nrow = 100)

for(i in 1:100){
  yrepbym2[i, ] <- rpois(357, posteriorbym2$mu[i, ]) # reparameterize
}

pp_check(y, yrepbym2, ppc_dens_overlay) + 
  theme_minimal() +
  xlim(0, 200) +
  labs(title = "Actual versus fitted") + xlim(0, 100)
```

This fit should also be improved once I finish variable selection.