### 2020-01-21 chapters/03_methods.Rmd
push new stan to run

---
title: "Methods"
author: "Raven McKnight"
date: "1/16/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Poisson Regression

Each model presented in this honors is built upon a simple Poisson regression. Poisson regressions are used to model count data and assume that the natural log of the mean of response variable $Y$ can be modeled by a linear combination of predictors. Poisson regressions are a form of generalized linear models with the natural log as its link function. For response variable $Y$ and covariates $x_1, x_2, ..., x_k$, the basic Poisson regression can be written

\begin{align}
\tag{1}
Y_i & \sim \text{Poisson}(E_i\lambda_i) \\
\eta_i  = log(\lambda_i) & = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k \\
\end{align}

where $\beta_0$ is an optional intercept term and $E_i$ is an offset term. The offset $E_i$ is often termed "exposure" in applications such as disease risk modeling where the offset may be the number of observed cases in a previous year, etc.

The most obvious assumption made in a Poisson regression is that response variable $Y$ follows a Poisson distribution. The primary assumption this makes about our data is that $\text{Var}(Y) = \text{E}(Y)$. When this assumption is not met, we can either use a generalization of the Poisson distribution, such as a Negative-Binomial distribution, or add a parameter to account for extra-Poisson variance.

For the purposes of this study, we can add a set of heterogeneous random effects $\theta$ to account for overdispersion (where Var($Y$) > E($Y$)). Following **cite mitzi morris**, the second line of **(1)** can be written

\begin{align}
\tag{2}
\eta_i = log(\lambda_i) & = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k + \theta \sigma
\end{align}

where $\theta$ and $\sigma$ are given normal priors centered around 0, such as $\theta \sim$ Normal(0, 1) and $\sigma \sim$ Normal(0, 5). The addition of these random effects improve fits on data with higher-than-expected variance.

# Horseshoe Priors and Variable Selection

Often, the coefficients $\beta_1, ..., \beta_k$ are given simple normal priors such as $\beta_k \sim$ Normal(0, 1) in the case of standardized parameters. When parameter vector $x$ contains many possible covariates, we may reasonably assume that some entries in $\beta$ are 0 -- in other words, that some of the parameters in $x$ have no effect on $Y$. In frequentist applications, we might use a method such as the LASSO (Least Absolute Shrinkage and Selection Operator) to identify relevant variables.

## Spike and Slab

In the Bayesian setting, there are two primary methods for variable selection. The first is called the "spike-and-slab" prior **(cite)** and has often been considered the "gold standard" for sparse Bayesian regression, or variable selection **(Vehtari)**. This prior is often expressed as

\begin{align}
\tag{3}
\beta_j | \lambda_j, c, \epsilon & \sim \lambda_j \text{Normal}(0, c^2) + (1-\lambda_j)\text{Normal}(0, \epsilon^2) \\
\lambda_j & \sim \text{Bernoulli}(\pi)
\end{align}

for $j = 1, 2, ..., J$ where $\lambda_j \in {0, 1}$ indicates whether $\beta_j$ is from the "spike" (ie, near zero) or the "slab" (ie, nonzero). Priors must be set for the width of the slab $c$ and the prior inclusion probability $\pi$. Here, $\pi$ reflects our prior understanding of the sparsity of $\beta$.

## Horseshoe

The second method for Bayesian sparse regression is to give $\beta$ some continuous *sparsity inducing prior*. **Van Erp et al** provide an introduction to many such continous priors. Ridge regression and the Bayesian LASSO are two of the most approachable priors, but all sparsity inducing priors utilize the same logic: give $\beta$ a prior with *most* of its mass near 0, to shrink irrelevant coefficients, and the rest of its mass "far" from 0. Identifying "far" depends more or less on the specific data and model in question.

The horseshoe prior, proposed by **Carvalho et all, 2010** is a popular choice in Bayesian literature, in part because it is similar to the "gold standard" spike-and-slab. Following the notation of **Vehtari & Pironeen**, the horseshoe prior can be expressed

\begin{align}
\tag{4}
\beta_k \text{ | } \lambda_k, \tau & \sim \text{Normal}(0, \tau^2\lambda_k^2) \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1)
\end{align}

for $k = 1, 2, ..., K$. The horseshoe prior is so-named because for fixed values $\tau = \lambda_k = 1$, the prior resemebles a Beta(1/2, 1/2) distribution, or a horseshoe.

The horseshoe is often favored because it is a global-local shrinkage prior. This means that $\tau$ shrinks *all* parameters towards 0 while the local parameter $\lambda_k$ and the heavy Cauchy tails allow larger coefficients to remain unshrunk. While this is precisely the goal of a sparsity inducing prior, the horseshoe prior can fail to regularize large coefficients *at all*, which can be a problem when parameters are weakly identified by data. Additionally, there is no consensus in the literature for assigning priors to $tau$. **Vehatari..** introduced the *regularized horseshoe* to address both of these shortcomings. The regularized horseshoe builds upon **(3)** such that

\begin{align}
\tag{5}
\beta_k \text{ | } \lambda_j, \tau, c & \sim \text{Normal}(0, \tau^2 \tilde{\lambda_k^2}) \\
\tilde{\lambda_k^2} & = \frac{c^2\lambda_k^2}{c^2 + \tau^2\lambda_k^2} \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1) \\
c & \sim ~ \text{Inverse-Gamma}(a, b)
\end{align}

where $c$ > 0 helps to regularize $\beta_k$ far from zero. When $c$ approaches infinity, the regularized horseshoe becomes the standard horseshoe. **This is a place where my understanding is weak** Using the spike-and-slab as intuition, c corresponds to the width of the slab. **Double check this, but I think** This is, in essence, how we describe "far" from 0.

# Spatial Structure

In general, we expect areas with