### 2021-01-26 2020-factbook-ideas.Rmd
reorganize 2019 & 2020 files

---
title: "2020 rider factbook ideas"
output:
  MetroTransitr::metro_html: default
  html_document:
    number_sections: no
    theme: paper
    toc: yes
    toc_float: yes
  pdf_document: default
---
<style type="text/css">
html {
  max-width: 10.5in;
  margin: 1.25in 0.75in;
  font-family: "Calibri", helvetica, sans-serif;
  font-size: 11pt;
}

body { width: 8.5in;
       max-width: 10.5in;
}
</style>

```{r setup, include=TRUE}
library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

```{r data, include = FALSE}
## 2019 data
dat19 <- readRDS("2019/data/modeling-dat/mod_dat_unscaled.RDS")
setDT(dat19)

dat20 <- readRDS("covid/data/modeling-dat/mod_dat_unscaled.RDS")
setDT(dat20)
```

```{r change}
setnames(dat20, "daily_boards", "daily_boards20")
boarding_change <- dat19[, .(GEOID, daily_boards)][dat20[, .(GEOID, daily_boards20)], on = "GEOID"]
boarding_change <- boarding_change[daily_boards != 0 & daily_boards20 != 0]
boarding_change[, change := daily_boards20 - daily_boards]
boarding_change[, perc_change := (change/daily_boards)*100]

avg_perc_change <- mean(boarding_change$perc_change, na.rm = TRUE)
```

```{r renters}
# predictors are the same in both 19 and 20

# get 75th percentile of renters?
#quantile(dat19$perc_rent, c(0.25, 0.5, 0.75, 0.90), na.rm = TRUE)
# 75th percentile rounds to 50% which is nice

renters75 <- dat19[perc_rent >= 0.50] # 447 bgs with ridership
renters50 <- dat19[perc_rent >= .21]
renters90 <- dat19[perc_rent >= .72]

avg_renters_perc_change <- mean(boarding_change[GEOID %in% renters75$GEOID]$perc_change, na.rm = TRUE)

avg_perc_change - avg_renters_perc_change
```

"Ridership in neighborhoods with a high percentage of renters decreased by ~9% less than the average block group"

## "Ridership in neighborhoods with at least 50% renters decreased by ~9% less than the average neighborhood."

high percentage = 75th percentile

```{r vehicles}
quantile(dat19$perc_no_veh, c(0.25, 0.5, 0.75, 0.90), na.rm = TRUE)

veh75 <- dat19[perc_no_veh >= 0.10] 
veh50 <- dat19[perc_no_veh >= .035]
veh90 <- dat19[perc_no_veh >= .21]

avg_veh_perc_change <- mean(boarding_change[GEOID %in% veh75$GEOID]$perc_change, na.rm = TRUE)

avg_perc_change - avg_veh_perc_change
```

"Ridership in neighborhoods with a high percentage of households with no vehicles decreased by ~9% less than the average block group"

## "Ridership in neighborhoods with at least 10% of households with no vehicle decreased by ~9% less than the average neighborhood"

```{r rail}
rail <- dat19[lightrail == 1]
avg_rail_perc_change <- mean(boarding_change[GEOID %in% rail$GEOID]$perc_change, na.rm = TRUE)

avg_perc_change 
avg_rail_perc_change
```
Only 2% less in block groups with rail stations